
```{r }
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
```{r }


## Loading and preprocessing the data
```{r}
library(ggplot2)
```
Download the zip file to the current working directory
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","datafile.zip")
```

Unzip the file
```{r}
unzip("datafile.zip")
```

Readfile
```{r}
readact <- read.csv("activity.csv", header = TRUE)
```


## What is mean total number of steps taken per day?
```{r}
tapply(readact$steps, readact$date, FUN=sum, na.rm = TRUE)
sumsteps <- tapply(readact$steps, readact$date, FUN=sum, na.rm = TRUE)
```
Make a histogram of the total number of steps taken each day
```{r}
hist(sumsteps,xlab = "Steps", ylab = "Date", main = "Total Steps Daily")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(sumsteps, na.rm = TRUE)
median(sumsteps, na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r}
aggregate(x=list(steps=readact$steps), by=list(interval=readact$interval),FUN=mean, na.rm = TRUE)
avgpat <- aggregate(x=list(steps=readact$steps), by=list(interval=readact$interval),FUN=mean, na.rm = TRUE)
```
Make a time series plot (type="l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)
```{r}
ggplot(data=avgpat, aes(x=interval, y=steps)) + geom_line() + xlab("5-minute interval") + ylab("average number of steps taken, averaged across all days")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avgpat[avgpat$steps==max(avgpat$steps),1]
```
## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
display the output
```{r}
sum(is.na(readact))
sumnas <- sum(is.na(readact))
```

Fillng missing values  with mean of 5-minute interval
```{r}
mvalues <- function(steps,interval){
NAval <- NA
if (!is.na(steps)) 
  NAVal <- c(steps)
else NAVal <- (avgpat[avgpat$interval==interval, "steps"])
return(NAVal)
}
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newdata <- readact
newdata$steps <- mapply(mvalues, newdata$steps,newdata$interval)
```

Make a histogram of the total number of steps taken each day 
```{r}
newdatatotsteps <- tapply(newdata$steps, newdata$date, FUN=sum)
hist(newdatatotsteps, xlab = "total number of steps taken each day")
```

Calculate and report the mean and median total number of steps taken per day.
```{r}
mean(newdatatotsteps)
median(newdatatotsteps)
```

                                                                        
Do these values differ from the estimates from the first part of the assignment? 
Answer is Yes.

What is the impact of imputing missing data on the estimates of the total daily number of steps?
Answer is the values are higher after imputing missing values.



## Are there differences in activity patterns between weekdays and weekends?
Answer is Yes.
For this part the weekdays() function may be of some help here.Use the dataset with the filled-in missing values for this part.
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
daytype <- function(date){
  weekday <- weekdays(date)
  if (weekday %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")) 
    return("Weekday")
  else if (weekday %in% c("Saturday","Sunday"))
    return("Weekend")
}
newdata$date <- as.Date(newdata$date)
newdata$day <- sapply(newdata$date,FUN=daytype)
```

Make a panel plot containing a time series plot (type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
avgwsteps <- aggregate(steps ~ interval + day, data = newdata, mean)
ggplot(avgwsteps, aes(interval,steps)) + geom_line() + facet_grid(day~.) + xlab("5-minute interval") + ylab("average number of steps across all weekday days or weekend days")
```
```

